---
title: "Combine UCLA Phonetics Lab Archive metadata with Glottolog metadata"
author: "Steven Moran"
date: "2023-10-09"
output: github_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

I took the the UCLA Phonetics Archive:

* http://archive.phonetics.ucla.edu/Language%20Indices/index_available.htm

And first created a CSV file of their language names and ISO codes by hand.

Then we load the hand created table.

```{r, warning=FALSE, message=FALSE}
ucla <- read_csv('ucla_metadata.csv')
```

Then load the Glottolog geo data.

```{r, warning=FALSE, message=FALSE}
glottolog_geo <- read_csv(url('https://cdstar.eva.mpg.de//bitstreams/EAEA0-B701-6328-C3E3-0/languages_and_dialects_geo.csv'))
```

Let's see what doesn't match by joining them. Actually, looks surprisingly pretty good -- only a few retired ISO codes that we can curate by hand, if needed.

```{r}
ucla$iso_6393 <- tolower(ucla$`Ethnologue Code`)
ucla_glottlog <- left_join(ucla, glottolog_geo, by=c("iso_6393"="isocodes"))
```

Now we can map them. :)

```{r}
ggplot(data=ucla_glottlog, aes(x=longitude, y=latitude)) + 
  borders("world", colour="gray50", fill="gray50") + 
  geom_point() +
  theme_bw()
```


Let's add in language family data. This must be downloaded from Glottolog (updated per release) instead of read directly from the website, like above.

```{r, warning=FALSE, message=FALSE}
glottolog_families <- read_csv('glottolog_languoid/languoid.csv')
```

Join them again.

```{r}
ucla_glottlog <- left_join(ucla_glottlog, glottolog_families)
```

And map them by family.

```{r}
ggplot(data=ucla_glottlog, aes(x=longitude, y=latitude, color=family_id)) + 
  borders("world", colour="gray50", fill="gray50") + 
  geom_point() +
  theme_bw() +
  theme(legend.position="none")
```

Write the combined data as csv.

```{r}
write_csv(ucla_glottlog, 'ucla_glottlog.csv')
```

